Precise Data Flow Analysis\\in the Presence of Correlated Method Calls\footnote{This research was supported by the Natural Sciences and Engineering Research Council of Canada and the Ontario Ministry of Research and Innovation.}

### 2015-06-19 sas2015_final/content/intro.tex
creating final version without comments and redundant files

\commentout{
Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs)
that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the CFG.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.
}

%Since a CFG is an over-approximation of the possible flows 
A control-flow graph (CFG) is an over-approximation of the possible flows 
of control in concrete executions of a program. It may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends on its ability to detect and disregard such infeasible paths.
The \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}
is a general data-flow analysis algorithm that 
%by  Reps, Horwitz, and Sagiv
%is a general data-flow analysis
%algorithm
%for computing solutions to finite distributive data-flow problems such 
%that can be instantiated to solve problems such
%as reaching definitions, available expressions, and taint analysis.
%A distinguishing 
%characteristic of IFDS is that it
avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. %Sagiv, Reps, and Horwitz
%also presented
The \textit{Interprocedural Distributive Environment} (IDE) algorithm~\cite{sagiv1996precise}
has the same property, but
%that similarly only considers properly matched call/return
%edges, but that 
supports a broader range of data-flow problems.
%by expanding the 
%domain of flow functions to \textit{environments} that go beyond the data-flow facts 
%considered by IFDS.

This paper presents an approach to data-flow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
%In such cases,
If the
method calls are polymorphic
(i.e., the method invoked depends on the run-time type of the receiver),
%(i.e., if they dispatch to different method definitions 
%depending on the type of the receiver expression at run time),
then their dispatch 
behaviours are correlated, and some of the paths between them are infeasible. A recent paper~\cite{DBLP:journals/scp/Tip15} made this observation
but did not present any concrete algorithm to take advantage of it.
%, and we are not aware 
%of any existing data-flow analysis that is capable of avoiding infeasible paths that arise 
%in the presence of correlated method calls.
% (see Section~\ref{sec:Related} for a discussion
%of related work).

%\input{figures/fig-transformation}

%The approach taken in our work is to 
Our approach transforms an IFDS problem into an
IDE problem that precisely accounts for infeasible paths due to correlated calls. 
The results of this IDE problem can be mapped back to the data-flow domain of the 
original IFDS problem, but are more precise than the results of directly applying
the IFDS algorithm to the original problem.
%
We present a formalization of the transformation and prove its correctness:
specifically, we prove it still soundly considers all paths that are feasible,
and that it avoids flow along all paths that are infeasible due to
correlated calls.
%as follows. 
%First, we derive an IDE problem that is equivalent to the original IFDS problem,
%as described by Reps et al.~\cite{sagiv1996precise}.
%Solving this ``Equivalence-IDE Problem'' with the IDE algorithm yields a result that
%is equivalent to the one obtained by applying the IFDS algorithm to the original IFDS problem.
%Then, we derive a ``Correlated-Calls 
%IDE Problem'' from the original IFDS problem and show 
%that the solution to this problem, when converted to an IFDS result, is more precise
%than the original IFDS result.
%We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
%This is illustrated schematically
%in Figure~\ref{fig:transformations}%
%\footnote{ 
%  The labels %$\resultIDE$, $\resultIFDS$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$
%  on edges in 
%  Figure~\ref{fig:transformations} reflect a number of mappings and %projections 
%  that will be defined in Section~\ref{sec:cca} and that can be %ignored here.
%}.

@@

%We also report on preliminary experiments in which our correlated-calls transformation
%is applied to an IFDS formulation of a simple taint analysis. Our results show that
%solving the resulting IDE problem avoids infeasible paths due to correlated calls as
%expected.
%
%In summary, the contributions of this paper are as follows:
%\begin{itemize}
%  \item
%    We present a general approach for transforming IFDS problems into corresponding
%    IDE problems that avoid infeasible paths due to correlated method calls and
%    prove its correctness. 
%  \item
%    We implemented the approach in Scala, on top of the WALA program analysis framework
%    and report on preliminary experiments. 
%\end{itemize}

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:cca} presents the correlated-calls transformation, 
states the correctness properties%
\footnote{
  Detailed proofs of our lemmas and theorems can be found in
  the \reportOrAppendix.
}, and discusses our implementation.
%
%Our implementation is discussed in Section~\ref{sec:Evaluation}.
%
Related work is discussed in Section~\ref{sec:Related}.
%
Finally, Section~\ref{sec:Conclusions} presents conclusions and directions for future work.

>>>>

A control-flow graph (CFG) is an over-approximation of the possible flows 
of control in concrete executions of a program. It may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends on its ability to detect and disregard such infeasible paths.
The \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}
is a general data-flow analysis algorithm that 
avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. 
The \textit{Interprocedural Distributive Environment} (IDE) algorithm~\cite{sagiv1996precise}
has the same property, but
supports a broader range of data-flow problems.

This paper presents an approach to data-flow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
If the
method calls are polymorphic
(i.e., the method invoked depends on the run-time type of the receiver),
then their dispatch 
behaviours are correlated, and some of the paths between them are infeasible. A recent paper~\cite{DBLP:journals/scp/Tip15} made this observation
but did not present any concrete algorithm to take advantage of it.

Our approach transforms an IFDS problem into an
IDE problem that precisely accounts for infeasible paths due to correlated calls. 
The results of this IDE problem can be mapped back to the data-flow domain of the 
original IFDS problem, but are more precise than the results of directly applying
the IFDS algorithm to the original problem.
We present a formalization of the transformation and prove its correctness:
specifically, we prove it still soundly considers all paths that are feasible,
and that it avoids flow along all paths that are infeasible due to
correlated calls.

@@

The remainder of this paper is organized as follows.
Section~\ref{sec:MotivatingExample} presents a motivating example.
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
Section~\ref{sec:cca} presents the correlated-calls transformation, 
states the correctness properties%
\footnote{
  Detailed proofs of our lemmas and theorems can be found in
  the \reportOrAppendix.
}, and discusses our implementation.
Related work is discussed in Section~\ref{sec:Related}.
Finally, Section~\ref{sec:Conclusions} presents conclusions and directions for future work.

### 2015-06-19 sas2015_final/content/related.tex
creating final version without comments and redundant files

\section{Related Work}
  \label{sec:Related}
%\otodo{Should we remove first initials in this section?}
The IFDS algorithm is an instance of the functional approach to data-flow analysis developed by Sharir and Pnueli~\cite{pnueli1981two}.
\commentout{
Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.
}
IFDS has been used to encode a variety of data-flow problems such as typestate analysis%
\commentout{(determining which operations can be performed on an object at a given program point)}~\cite{naeem2008typestate,DBLP:conf/pldi/ZhangMGNY14} and shape analysis%
\commentout{(detecting errors and validating properties of programs at compile time)}%
~\cite{DBLP:conf/birthday/KreikerRRSWY13}. IFDS has been used~\cite{DBLP:conf/pldi/ArztRFBBKTOM14,tripp2009taj} and extended~\cite{DBLP:conf/sigsoft/LerchHBM14} to solve taint-analysis problems.

%The IFDS algorithm is implemented in two popular static-analysis frameworks for Java bytecode, the T.J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

\commentout{
WALA is a framework for static analysis on Java bytecode developed by the IBM T.J.~Watson Research Center.
In the implementation of our work, we use WALA to build and traverse the supergraph (a special kind of control-flow graph) of a Java program\footnote{However, we do not use WALA's IFDS implementation, as explained in~Section~\ref{chapter:eval}.}.

Soot is a framework for program analysis and optimization on Java bytecode, developed by the Sable Research Group at McGill University. Unlike WALA, Soot also has an implementation of the IDE algorithm. The IFDS and IDE implementations for Soot are part of the Heros project~\cite{bodden2012inter}.

Whereas one advantage of Soot's IFDS implementation (and other static analysis tools) is ease of use and extensibility, WALA's primary focus is efficiency. For example, WALA uses bit-vectors to represent some of the analysis data types, like local variables and parameters. Another difference is that WALA's intermediate representation of a program uses static single assignment (SSA) form~\cite{cytron1991efficiently}. SSA form is a representation of the program in which each variable has only one definition (assignment). SSA can make data-flow analysis simpler and more efficient~\cite{appel1998ssa}.
}

@@

\commentout{
Work on improving the IFDS algorithm includes the Practical Extensions to the IFDS algorithm~\cite{naeem2010pei}.
\commentout{The paper presents four extensions to the IFDS algorithm.}
Two of the four extensions improve the efficiency of the IFDS analysis for certain classes of IFDS problems. Another extension widens the class of problems applicable for the IFDS analysis. 
\commentout{However, those extensions do not affect the precision of IFDS problems.}
Our analysis, in contrast, does not improve the efficiency or generality of IFDS, but it allows us to solve IFDS problems more precisely.
The fourth extension is targeted towards programs that are represented in SSA form. Executing the IFDS analysis on such programs results in loss of precision in the presence of control-flow constructs (e.g. conditionals and loops), compared to programs in non-SSA form.
The extension makes the IFDS analysis on programs in SSA form as precise as on programs that are not represented in SSA form. In contrast, the correlated-calls analysis is applicable to programs in both SSA and non-SSA forms. Even if applied to a program in SSA form, our analysis and the extension improve the precision of IFDS in unrelated situations: the first analysis handles correlated calls, and the latter handles control-flow constructs. Thus, an IFDS analysis could benefit from both precision improvements independently.
}

Bodden et~al.~\cite{bodden2013spl} presents a framework
for applying IFDS analyses to software product lines. Their approach enables the analysis of
all possible products derived from a product line in a single analysis pass. Like our approach,
their approach transforms IFDS problems to IDE problems. The micro-functions keep track
of the possible program variations specified by the product line.
Rodriguez and Lhot\'ak evaluate a parallelized implementation of the IFDS algorithm using actors~\cite{rodriguez2010concurrent} that can take advantage of multiple processors.
%However, neither of those works is concerned with improving the precision of IFDS results.

\commentout{
The correlated-calls analysis improves the precision of a data-flow analysis by eliminating a special type of infeasible paths. This is similar to the idea of context-sensitive analysis: just as a context-sensitive analysis eliminates infeasible paths from the end of a procedure to the call sites that do not match the given procedure call, the correlated-calls analysis eliminates infeasible paths caused by correlated method calls.
}

@@

Recent work on correlation tracking for JavaScript~\cite{DBLP:conf/ecoop/SridharanDCST12} 
also eliminates infeasible paths. Instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
of an object. 
%Unlike in more static languages,
%property names in JavaScript are first class values (strings). When a loop indexed by
%property names updates the properties of objects, the loop body is extracted into a function
%and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.

\commentout{
The idea of eliminating infeasible paths caused by correlated calls is similar to M.~Sridharan et~al.'s work on improving the precision of pointer analysis for JavaScript programs~\cite{DBLP:conf/ecoop/SridharanDCST12}. For each pointer, a pointer analysis determines the possible set of objects (the \textit{points-to} set) that the pointer can reference at a given program point. In JavaScript, it is challenging to compute the points-to set of fields because in general, field names can be derived from arbitrary expressions and bound at runtime.
As a result, an imprecise data-flow analysis will include infeasible paths between values of the form \verb'o[p]' (access of a property \verb'p' of object \verb'o'), where at compile time, \verb'p' can be bound to different values.
The idea of the paper is to track all dynamic property accesses (reads and writes) on an object \verb'o' with property name \verb'p'. The code snippets containing the references \verb'o[p]' are then extracted into a separate function $f$. The analysis is then run so that for each possible value of \verb'p', $f$ is analyzed separately; therefore, for a given property name, all correlated objects with that name are analyzed together.

The differences between this method of tracking correlated calls and our analysis are the following.
\begin{itemize}
  \item \textit{Type of target data-flow analysis} whose precision is to be improved. Our analysis improves the precision of IFDS data-flow analyses, whereas the JavaScript analysis improves the precision of pointer analysis.
  \item \textit{Target language}. Our analysis is for object-oriented languages where polymorphic methods, and not property names (which are known at compile time), cause infeasible paths.
  \item \textit{Different handling of correlated calls}. Extracting code that contains correlated calls into separate methods would not prevent infeasible paths. Instead, our analysis uses IDE flow functions to detect and eliminate infeasible paths caused by correlated calls.
\end{itemize}
}

>>>>

\section{Related Work}
  \label{sec:Related}
The IFDS algorithm is an instance of the functional approach to data-flow analysis developed by Sharir and Pnueli~\cite{pnueli1981two}.
IFDS has been used to encode a variety of data-flow problems such as typestate analysis~\cite{naeem2008typestate,DBLP:conf/pldi/ZhangMGNY14} and shape analysis%
~\cite{DBLP:conf/birthday/KreikerRRSWY13}. IFDS has been used~\cite{DBLP:conf/pldi/ArztRFBBKTOM14,tripp2009taj} and extended~\cite{DBLP:conf/sigsoft/LerchHBM14} to solve taint-analysis problems.

@@

Bodden et~al.~\cite{bodden2013spl} presents a framework
for applying IFDS analyses to software product lines. Their approach enables the analysis of
all possible products derived from a product line in a single analysis pass. Like our approach,
their approach transforms IFDS problems to IDE problems. The micro-functions keep track
of the possible program variations specified by the product line.
Rodriguez and Lhot\'ak evaluate a parallelized implementation of the IFDS algorithm using actors~\cite{rodriguez2010concurrent} that can take advantage of multiple processors.

@@

Recent work on correlation tracking for JavaScript~\cite{DBLP:conf/ecoop/SridharanDCST12} 
also eliminates infeasible paths. Instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
of an object. 
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.
